### 2016-12-11 doc/main.tex
refactored docs

\documentclass{bioinfo}
\copyrightyear{2016} \pubyear{2016}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Application Note}

\usepackage{url}
\begin{document}

\firstpage{1}

\history{Received on July 23, 2016; revised on October 30, 2016}
\editor{Associate Editor: Alfonso Valencia}

\subtitle{Application Note}
\title[Historian: accurate reconstruction of sequences and rates]{Historian: accurate reconstruction of ancestral sequences and evolutionary rates}
\author[Ian Holmes]{Ian Holmes$^{\text{\sfb 1}*}$}
\address{$^{\text{\sf 1}}$Department of Bioengineering, University of California, Berkeley, 94720, USA.}
\corresp{$^\ast$To whom correspondence should be addressed.}

\input{paper.tex}

\end{document}